# Reproducible Research: Peer Assessment 1

The data for this assignment consists of activity measurements:

* _steps_: Number of steps taking in a 5-minute interval (missing values are coded as NA).
* _date_: The date on which the measurement was taken in YYYY-MM-DD format.
* _interval_: Identifier for the 5-minute interval in which measurement was taken.

It is contained in the activity.zip file in this repo.

This knitr document has been run with the following global commands:

* No scientific notation for long numbers.
* No digits after the point for floating point numbers.

```{r setoptions, echo=TRUE}
options(scipen = 1, digits = 0)
```

## Loading and preprocessing the data

First, all library includes are listed. This way, it is immediately signalled
if all necessary packages are installed or not.

```{r, echo=TRUE}
library(plyr)
```

Then, the activity data is unzipped and read it into a data frame.

```{r, echo=TRUE} 
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

The step amounts in the activity data are aggregated by day. 

```{r, echo=TRUE}
activity_by_day <- ddply(activity, .(date), summarize, steps=sum(steps, na.rm=TRUE))
```

The figure below shows the histogram of steps taken by day.

```{r, echo=TRUE}
hist(activity_by_day$steps, 
     xlab="Steps", 
     ylab="Days", 
     main="Steps distribution (by day)")

mean <- mean(activity_by_day$steps)
median <- median(activity_by_day$steps)
```

On average, `r mean` steps were taken by day, with a median of `r median` 
steps.

## What is the average daily activity pattern?

The step amounts are aggregated by (5-minute) interval and averaged over all
days.

```{r, echo=TRUE}
activity_by_interval <- ddply(activity, .(interval), summarize, avg_steps=mean(steps, na.rm=TRUE))
```

The figure below shows a time series of average daily step pattern in 5-minute
intervals.

```{r, echo=TRUE}
plot(activity_by_interval$avg_steps, 
     type="l", 
     xaxt="n", 
     xlab="Time of day", 
     ylab="Average steps", 
     main="Daily activity pattern")
axis(1, 
     at=1:length(activity_by_interval$interval), 
     labels=activity_by_interval$interval)

max_steps_in_day <- activity_by_interval[which.max(activity_by_interval$avg_steps),]
```

The most steps in a single 5-minute interval on average were at `r max_steps_in_day$interval`.

## Imputing missing values

```{r, echo=TRUE}
valid <- complete.cases(activity$steps)
nr_of_NAs <- sum(!valid)
```

The activity data set contains `r nr_of_NAs` rows with NA values.

Missing values are imputed by taking the daily average steps for that interval
over the entire dataset.

```{r, echo=TRUE}
imputed_activity <- activity
for (i in 1:length(imputed_activity$steps)) {
    if (is.na(imputed_activity$steps[i])) {
        interval <- imputed_activity$interval[i] 
        index <- activity_by_interval$interval == interval
        imputed_activity$steps[i] <- activity_by_interval$avg_steps[index]
    }
}
```

The step amounts of the new imputed activity data are again aggregated by day. 

```{r, echo=TRUE}
imputed_activity_by_day <- ddply(imputed_activity, .(date), summarize, steps=sum(steps, na.rm=TRUE))
```

The histogram below shows steps taken by day of the imputed activity data.

```{r, echo=TRUE}
hist(imputed_activity_by_day$steps, 
     xlab="Steps", 
     ylab="Days", 
     main="Steps distribution (by day)")

imputed_mean <- mean(imputed_activity_by_day$steps)
imputed_median <- median(imputed_activity_by_day$steps)
```

The average by day of the imputed data is `r imputed_mean` steps, with a median
of `r imputed_median` steps.  
These values differ from the base activity data. The choice of imputing by mean
steps over all days by 5-minute interval combined with the fact that, when
missing values (NAs) are present, they are present for a full day has resulted
in the same mean and median for the imputed activity dataset. They are also
higher than in the base activity data set.

## Are there differences in activity patterns between weekdays and weekends?
