# Reproducible Research: Peer Assessment 1

The data for this assignment consists of activity measurements:

* _steps_: Number of steps taking in a 5-minute interval (missing values are coded as NA).
* _date_: The date on which the measurement was taken in YYYY-MM-DD format.
* _interval_: Identifier for the 5-minute interval in which measurement was taken.

It is contained in the activity.zip file in this repo.

This knitr document has been run with the following global commands:

```{r setoptions, echo=TRUE}
options(scipen = 1, digits = 1)
```

## Loading and preprocessing the data

First, all library includes are listed. This way, it is immediately signalled
if all necessary packages are installed or not.

```{r, echo=TRUE}
library(plyr)
```

Then, the activity data is unzipped and read it into a data frame.

```{r, echo=TRUE} 
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

The step amounts in the activity data are aggregated by day. 

```{r, echo=TRUE}
activity_by_day <- ddply(activity, .(date), summarize, steps=sum(steps))
```

The figure below shows the histogram of steps taken by day.

```{r, echo=TRUE}
hist(activity_by_day$steps, xlab="Steps", ylab="Days", main="Steps distribution (by day)")
mean <- mean(activity_by_day$steps, na.rm=TRUE)
median <- median(activity_by_day$steps, na.rm=TRUE)
```

On average, `r mean` steps were taken by day, with a median of `r median` 
steps.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
