# Reproducible Research: Peer Assessment 1

The data for this assignment consists of activity measurements:

* _steps_: Number of steps taking in a 5-minute interval (missing values are coded as NA).
* _date_: The date on which the measurement was taken in YYYY-MM-DD format.
* _interval_: Identifier for the 5-minute interval in which measurement was taken.

It is contained in the activity.zip file in this repo.

This knitr document has been run with the following global commands:

```{r setoptions, echo=TRUE}
options(scipen = 1, digits = 1)
```

## Loading and preprocessing the data

First, all library includes are listed. This way, it is immediately signalled
if all necessary packages are installed or not.

```{r, echo=TRUE}
library(plyr)
```

Then, the activity data is unzipped and read it into a data frame.

```{r, echo=TRUE} 
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

The step amounts in the activity data are aggregated by day. 

```{r, echo=TRUE}
activity_by_day <- ddply(activity, .(date), summarize, total_steps=sum(steps, na.rm=TRUE))
```

The figure below shows the histogram of steps taken by day.

```{r, echo=TRUE}
hist(activity_by_day$total_steps, xlab="Steps", ylab="Days", main="Steps distribution (by day)")

mean <- mean(activity_by_day$total_steps)
median <- median(activity_by_day$total_steps)
```

On average, `r mean` steps were taken by day, with a median of `r median` 
steps.

## What is the average daily activity pattern?

The step amounts are aggregated by (5-minute) interval and averaged over all
days.

```{r, echo=TRUE}
activity_by_interval <- ddply(activity, .(interval), summarize, average_steps=mean(steps, na.rm=TRUE))
```

The figure below shows a time series of average daily step pattern in 5-minute
intervals.

```{r, echo=TRUE}
plot(activity_by_interval$average_steps, type="l", xaxt="n", xlab="Time of day", ylab="Average steps", main="Daily activity pattern")
axis(1, at=1:length(activity_by_interval$interval), labels=activity_by_interval$interval)

max_steps_in_day <- activity_by_interval[which.max(activity_by_interval$average_steps),]
```

The most steps in a single 5-minute interval on average were at `r max_steps_in_day$interval`.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
